---
title: "Case Study Template (Rough Draft)"
output: html_document
date: "2026-07-28"
params:
  gageid: "02056900"
  startdates: !r as.Date(c("2008-06-11", "2007-06-10", "2002-05-01", "1986-05-12"))
  enddate: "2026-08-28"
  iterative_AGWRC: 0.97
  digits: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
basepath <- "/var/www/R"
source("/var/www/R/config.R")
```

```{r loop, warning = FALSE, echo = FALSE, message = FALSE}
library(knitr)
library(flextable)
library(dplyr)
library(ggplot2)

# create gage object environment for specified gage
gage_obj <- hydrotools::WaterGageDaily$new(gage_id = params$gageid,
                                           ds_in = ds)
# get gage data
all_data <- gage_obj$baseflow_workflow_data(omsite = omsite)
agwrc <- gage_obj$agwrc_fun(TRUE)

# Create forecast graphs loop for the given start dates
forecast_plot <- list()
forecast_table <- list()

for (i in seq_along(params$startdates)) {
  
  sd <- params$startdates[i]
  
  # Define different AGWRC forecast methods
  forecast_data <- gage_obj$baseflow_forecast(start_date = sd,
                                            AGWRC = "lm_constant",
                                            use_limits = TRUE)
  
  Iterative_AGWRC <- gage_obj$baseflow_forecast(start_date = sd, 
                                           AGWRC = params$iterative_AGWRC) %>%
    rename(Iterative_AGWRC = AGWRC) %>%
    rename(Itr_AGWRC_Flow = Forecast)
  
  Constant_AGWRC <- gage_obj$baseflow_forecast(start_date = sd, 
                                     AGWRC = "lm_constant") %>%
    rename(Constant_AGWRC = AGWRC) %>%
    rename(Con_AGWRC_Flow = Forecast)
  
  Variable_AGWRC <- gage_obj$baseflow_forecast(start_date = sd,
                                     AGWRC = "lm_variable") %>%
    rename(Variable_AGWRC = AGWRC) %>%
    rename(Var_AGWRC_Flow = Forecast)
  
  # Plot the forecast data
  forecast_plot[[as.character(sd)]] <- gage_obj$plot_baseflow_forecast(
  start_date = sd,
  forecast_days = 0:90,
  AGWRC = list(Constant_AGWRC = "lm_constant",
               Variable_AGWRC = "lm_variable",
               Iterative_AGWRC = params$iterative_AGWRC),
  return_plotly = FALSE,
  include_days_before = 60,
  use_limits = TRUE
)
  
  forecast_plot[[as.character(sd)]] <- forecast_plot[[as.character(sd)]] + 
    theme_minimal() +
    labs(
      x = "Date",
      y = "Flow (cfs)",
      title = "Observed and Projected Flows"
    )
  
# create forecast table
all_flow <- forecast_data %>%
  full_join(Constant_AGWRC, by = c("Day", "Date", "obs_flow" )) %>%
  full_join(Iterative_AGWRC, by = c("Day", "Date", "obs_flow" )) %>%
  full_join(Variable_AGWRC, by = c("Day", "Date", "obs_flow" ))

final_flow_df <- all_flow %>%
  select(Day, Date, obs_flow, Con_AGWRC_Flow, Var_AGWRC_Flow, Itr_AGWRC_Flow) %>%
  rename(Obs_Flow = obs_flow)

Forecast_df <- final_flow_df %>%
  filter(Day %in% c(15, 30, 45, 90)) %>% 
  mutate(across(c(Obs_Flow, Con_AGWRC_Flow, Var_AGWRC_Flow, Itr_AGWRC_Flow),
  ~ round(.x, params$digits)))

forecast_table[[as.character(sd)]] <- flextable(Forecast_df) %>%
  autofit()

}

```

```{r case study data, warning = FALSE, echo = FALSE, message = FALSE, results = "hide"}
library(dplyr)

# load in case study data from database

gage_obj$load_wshd_feat()

gage_props <- gage_obj$get_model_or_scenario_props(model_prop_code = "AGWRC-1.0",
                                                   include_proptext = TRUE)

case_studies <- gage_obj$get_model_or_scenario_props(model_prop_code = "AGWRC-1.0",
                                                     scenario_prop_code = "case_studies",
                                                     include_proptext = TRUE)

# convert the epoch start dates to yyyy-mm-dd format

cs_df <- case_studies %>%
  rename(epoch_startdate = startdate) %>%
  rename(epoch_enddate = enddate)

cs_df$startdate <- as.Date(as.POSIXct(cs_df$epoch_startdate,
                                             origin = "1970-01-01",
                                             tz = "UTC"))


cs_df$enddate <- as.Date(as.POSIXct(cs_df$epoch_enddate,
                                             origin = "1970-01-01",
                                             tz = "UTC"))

# extract just the case study conclusion text from gage_props

cs_conclusion <- gage_props$proptext[!is.na(gage_props$proptext)]

```

## Overview and Recommended AGWRC

AGWRC analysis for USGS **`r params$gageid`: `r gage_obj$gage_data_sf$monitoring_location_name`**. `r cs_conclusion` 

## Terminology

**Regression AGWRC**: the AGWRC calculated by imputing starting flow from manually selected flow event into AGWRC vs Flow Regression

**Event AGWRC**: the AGWRC from a known baseflow event identified by the workflow

**Iterative AGWRC**: the manually selected AGWRC that provides the "best fit" for the manually selected baseflow event

**Starting flow**: the observed flow at the first day of the manually selected baseflow event

**MAE**: Mean Absolute Error

**RMSE**: Root Mean Square Error

**MAPE**: Mean Absolute Percentage Error

## Regression Analysis

```{r regression plot, warning = FALSE, echo = FALSE}
gage_obj$plot_baseflow_agwrc(CI = TRUE, omsite = omsite, return_plotly = FALSE)

# R-squared interpretation based on value

rsqr <- all_data$lm_df$Rsq

interpret_rsqr <- function(rsqr) {
  if (rsqr < 0.01) {
    "Very little of the observed variation in the data can be   
    explained by the regression model. Historically, the regression has not performed   
    well at other gages with similar R-squared values."
    } else if (rsqr < 0.1) {
      "Little of the observed variation in the data can be 
      explained by the regression model. Historically, the regression has not consistenty
      performed well at other gages with similar R-squared values."
      } else if (rsqr < 0.2) {
        "Some of the observed variation in the data can be 
        explained by the regression model. Historically, the regression has performed 
        decently well at other gages with similar R-squared values."
        } else if (rsqr < 0.35) {
          "A considerable amount of the observed variation in the 
          data can be explained by the regression model. Historically, the regression has
          performed well at other gages with similar R-squared values."
          } else {
            "An exceptional amount of the observed variation in the
            data can be explained by the regression model (relative to other gages). 
            Historically, the regression has performed well at other gages with similar 
            R-squared values."
          }
}

rsqr_interpretation <- interpret_rsqr(rsqr)

```

**Image 1.** AGWRC vs Flow Regression

The R-squared value is: `r all_data$lm_df$Rsq`. `r rsqr_interpretation`

## Annual Low Flow Plot

```{r low flow plot, warning = FALSE, echo = FALSE, message = FALSE}
gage_obj$plot_low_flows()
```

**Image 2.** Annual low flow plot for 1-day, 7-day, 30-day, and 90-day low flows

## Baseflow Analysis

### First Baseflow Event Investigation

```{r baseflow plot 1, warning = FALSE, echo = FALSE, message = FALSE}
print(forecast_plot[[1]])
```

**Image 2.** Observed and Projected Flows from `r params$startdates[1]`

Although none of the AGWRC values were very effective in surviving the storms and catching the "troughs" following the peaks in flow, the iterative AGWRC performed the best of the three.

**Table 1.** Constant, Variable, and Iterative AGWRC Flow Forecast from `r params$startdates[1]`

```{r table 1, warning = FALSE, echo = FALSE, message = FALSE}
library(knitr)
library(flextable)
forecast_table[[1]]
```

### Second Baseflow Event Investigation

```{r baseflow plot 2, warning = FALSE, echo = FALSE, message = FALSE}
print(forecast_plot[[2]])
```

**Image 3.** Observed and Projected Flows from `r params$startdates[2]`

The iterative AGWRC overpredicted the first trough, but predicted the subsequent three troughs well. The iterative and variable AGWRC predictions were effective in surviving the first storm and predicting the first trough, but they greatly underpredicted the remaining flows.

**Table 2.** Constant, Variable, and Iterative AGWRC Flow Forecast from `r params$startdates[2]`

```{r table 2, warning = FALSE, echo = FALSE, message = FALSE}
library(knitr)
library(flextable)
forecast_table[[2]]
```

### Third Baseflow Event Investigation

```{r baseflow plot3, warning = FALSE, echo = FALSE, message = FALSE}
print(forecast_plot[[3]])
```

**Image 4.** Observed and Projected Flows from `r params$startdates[3]`

The iterative AGWRC was effective in the predicting the first trough, but overpredicted flow in the second trough where the flow approached 0 cfs. The iterative AGWRC and constant AGWRC produced very similar predictions and also overpredicted flow in mid August.

**Table 3.** Constant, Variable, and Iterative AGWRC Flow Forecast from `r params$startdates[3]`

```{r table 3, warning = FALSE, echo = FALSE, message = FALSE}
forecast_table[[3]]
```

### Fourth Baseflow Event Investigation

```{r baseflow plot 4, warning = FALSE, echo = FALSE, message = FALSE}
print(forecast_plot[[4]])
```

**Image 5.** Observed and Projected Flows from `r params$startdates[4]`

Of the three AGWRC predictions, the iterative AGWRC was the most effective in surviving the storms. The precipitation events at the end of July and the start of August impacted the accuracy of the 90-day prediction.

**Table 4.** Constant, Variable, and Iterative AGWRC Flow Forecast from `r params$startdates[4]`

```{r table 4, warning = FALSE, echo = FALSE, message = FALSE}
library(knitr)
library(flextable)
forecast_table[[4]]
```

```{r 4 plots w/ patchwork, warning = FALSE, echo = FALSE, message = FALSE}
library(ggplot2)
library(patchwork)

if (length(forecast_plot) == 4) {
forecast_plot <- lapply(forecast_plot, function(p) {
  p + labs(title = NULL)
})

(forecast_plot[[1]] | forecast_plot[[2]]) /
  (forecast_plot[[3]] | forecast_plot[[4]]) +
  plot_annotation(
    title = "Forecast Comparison") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5, face = "bold", size = 16)
  )

} else {
  message("Forecast comparison plot skipped: need 4 forecasts, found ", length(forecast_plot))
}
```

**Image 5.** Observed and Projected Flows for All Four Start Dates

## Prediction Accuracy Analysis

### MAE, RMSE, MAPE, Bias, and $R^2$ Analysis 

**Table 5.** Calculations of the MAE, RMSE, MAPE, Bias, and $R^2$ values for all group IDs for **`r params$gageid`: `r gage_obj$gage_data_sf$monitoring_location_name`**

```{r error calculations, warning = FALSE, echo = FALSE, message = FALSE}
devtools::source_url("https://raw.githubusercontent.com/HARPgroup/baseflow_storage/refs/heads/Emily_bf_workflow/PredictionAccuracyUpdated.R")

flow_csv <- gage_obj$gage_data |>
 select(time, value) |>
 rename(Date = time, Observed = value)

baseflow_csv <- all_data$trimmed_events_df |>
  mutate(Date = as.Date(Date)) |>
  group_by(GroupID) |>
  mutate(duration = as.numeric(max(Date) - min(Date))) |>
  slice(1) |>
  ungroup() |>
  filter(duration >= 7) |>
  select(GroupID, Date, Flow, AGWRC, duration)

regression_csv <- all_data$lm_df

m <- regression_csv$m
b <- regression_csv$b

Results <- PredictionAccuracy(flow_csv, baseflow_csv, days = c(0:15), AGWRC = "lm_constant", m, b)

Results$summary_table <- Results$summary_table %>%
  colformat_double(
    j = c("MAPE", "MAE", "RMSE", "Bias", "r2"),
    digits = params$digits
  )
Results$summary_table
```

**Table 6.** All GroupIDs with an MAPE over 20%

```{r high MAPE, warning = FALSE, echo = FALSE, message = FALSE}
filtered_group_stats <- Results$group_stats %>%
  filter(MAPE > 20) %>% 
   mutate(across(c(MAPE, MAE, RMSE, Bias, r2),~round(.x, params$digits)))

flextable(filtered_group_stats) %>%
  theme_vanilla() %>%
  autofit()
```

The following GroupIDs in table 6 present a list of event that did not preform well and should be treated with caution.

### Minium Flow Forecast Accuracy

Determines accurate of a forecast. More text to come as function is changed.

```{r accuracy analysis, warning = FALSE, echo = FALSE, message = FALSE, include=FALSE}
# library(zoo)
# 
# devtools::source_url("https://raw.githubusercontent.com/HARPgroup/baseflow_storage/refs/heads/willv_bf_workflow/workspace/will_videll/min_flow_forecast_accuracy.R")
# 
# min_flow_df <- min_flow_accuracy(gage_obj, event_df = all_data$trimmed_events_df, AGWRC = c("lm_variable"))
# 
# min_flow_df$eventdf1
# 
# plot_gage_event_minima <- lapply(params$startdates, function(start_date) {
# 
#   plot_event_minima(
#    gage_obj,
#    start_date
#    )
# 
#  })
```

```{r first minima plot, warning = FALSE, echo = FALSE, message = FALSE}
# plot_gage_event_minima[[1]]
```

**Image 6.** Observed and Projected Flows from `r params$startdates[1]`

```{r second minima plot, warning = FALSE, echo = FALSE, message = FALSE}
# plot_gage_event_minima[[2]]
```

**Image 7.** Observed and Projected Flows from `r params$startdates[2]`

```{r third minima plot, warning = FALSE, echo = FALSE, message = FALSE}
# plot_gage_event_minima[[3]]
```

**Image 8.** Observed and Projected Flows from `r params$startdates[3]`

```{r fourth minima plot, warning = FALSE, echo = FALSE, message = FALSE}
# plot_gage_event_minima[[4]]
```

**Image 9.** Observed and Projected Flows from `r params$startdates[4]`

## Case Studies

```{r case study graph, warning = FALSE, echo = FALSE, message = FALSE}
library(ggplot2)
library(stringr)

cs_plot <- list()

for(i in 1:nrow(cs_df)) {
  
  cs_sd <- cs_df$startdate[i]
  
  BPJ_AGWRC <- as.numeric(cs_df$propcode[i])
  
  cs_text <- str_wrap(cs_df$proptext[i])
  
  p_label <- paste0("Interpretation: ", cs_text)
  
  p <- gage_obj$plot_baseflow_forecast(
  start_date = cs_sd,
  forecast_days = 0:90,
  AGWRC = list(Constant_AGWRC = "lm_constant",
               Variable_AGWRC = "lm_variable",
               Iterative_AGWRC = BPJ_AGWRC),
  return_plotly = FALSE,
  include_days_before = 60,
  use_limits = TRUE) + 
    labs(caption = cs_text) +
    theme(plot.caption = element_text(hjust = 0.5)) +
  scale_color_discrete(
    name = NULL,
    labels = c(
      "Constant_AGWRC" = "Constant AGWRC",
      "Variable_AGWRC" = "Variable AGWRC",
      "Iterative_AGWRC" = paste0("BPJ AGWRC = ", BPJ_AGWRC)
    )
  )
  
  cs_plot[[i]] <- p
  
  print(p)
}

```

